Demes: a specification for interchanging complex demography

### 2021-06-21 paper.tex
Updates from review.

\begin{figure}
\begin{minipage}{0.48\textwidth}
\begin{tcolorbox}
\inputminted[fontsize=\scriptsize,linenos,numbersep=5pt]{yaml}{models/IM.yaml}
\end{tcolorbox}
\begin{tcolorbox}
\includegraphics[width=\textwidth]{fig/IM}
\end{tcolorbox}
\end{minipage}\hfill
\begin{minipage}{0.48\textwidth}
\begin{tcolorbox}
\inputminted[fontsize=\scriptsize,linenos,numbersep=5pt]{yaml}{models/IM-resolved.yaml}
\end{tcolorbox}
\end{minipage}\\
\caption{\label{fig-example}
Example isolation-with-migration Demes model. (A) The compact non-redundant
representation expressed using YAML. (B) The same model in the fully-qualified
form. (C) A visual representation of the model using \texttt{demesdraw}.
The YAML descriptions in (A) and (B) correspond exactly to a JSON description,
but are much more concise.
}
\end{figure}

@@

The high-level description of Demes models is focused on the efficiency of
human understanding and the avoidance of errors. We have adopted the widely
used YAML format~\citep{ben2009yaml} as the recommended means of interchanging
Demes models (see Figure~\ref{fig-example} for exampels). YAML is a data
serialisation language with an emphasis on simplicity, and interoperates well
with JSON (indeed, YAML 1.2 is a superset of JSON). We chose YAML over JSON
because although JSON is an excellent format for data interchange, it is
ill-suited for human understanding or manipulation. We also considered other
data formats such as TOML, but chose YAML because of its greater popularity and
software support. Since the Demes data model is defined in JSON Schema,
however, there is no formal dependency on YAML and implementations may choose
to use JSON directly if they wish (e.g., for greater efficiency).

@@

The Demes specification is by design simple and static---we wish to
unambiguously describe a demographic model with fully specified
and fixed parameters. This simplicity means that we cannot directly
specify distributions or estimated confidence intervals
for those parameters. While it is not difficult to imagine extending
the specification in ways that would allows us to do this, it would
greatly increase the complexity of parsers and would likely be of
little practical use in any case. This is because the parameters of
demographic models are typically tightly coupled, and cases in which
distributions for different parameters can be simply described are rare.
In this situation, the simplest way to describe estimated
distribution is simply to list a large number samples from
the posterior. While writing out a large number of Demes models in
YAML format may seem inefficient, it can in fact be a compact
way to describe these distributions.
\jkcomment{TODO: add in a quick example here where we take an IM model
and simulate a posterior distribution under some reasonable model. Report
how much space it takes to represent a large number of samples from this
distribution}

Similarly, one might be interested in running simulations in which
the demographic model parameters are drawn from a distribution, e.g.,
in ABC inference~\citep{beaumont2002approximate}. Allowing for
the many distributions of interest would greatly increase the complexity
parsers, with a relatively limited benefit. It would be far simpler
and more powerful to write some code to sample directly from the
distributions of interest, using, for example the Demes Python
API.
\jkcomment{This is garbled, but we get the point? You're probably never
going to be doing ABC from a completely general Demes model as input.
There will be some basic model that you have in mind, that you're generating
distributions of parameters around. So, what's the point in writing
general machinery to do this sampling, when you could write it yourself in
10 lines of Python?}

>>>>

\begin{figure}
\begin{minipage}{0.48\textwidth}
\begin{tcolorbox}
\inputminted[fontsize=\scriptsize,linenos,numbersep=5pt]{yaml}{models/IM.yaml}
\end{tcolorbox}
\begin{tcolorbox}
\includegraphics[width=\textwidth]{fig/IM}
\end{tcolorbox}
\end{minipage}\hfill
\begin{minipage}{0.48\textwidth}
\begin{tcolorbox}
\inputminted[fontsize=\scriptsize,linenos,numbersep=5pt]{yaml}{models/IM-resolved.yaml}
\end{tcolorbox}
\end{minipage}\\
\caption{\label{fig-example}
Example isolation-with-migration Demes model. (A) The compact non-redundant
representation expressed using YAML. (B) The same model in the fully-qualified
form. (C) A visual representation of the model using \texttt{demesdraw}.
The YAML descriptions in (A) and (B) correspond exactly to a JSON description,
but are much more human-readable.
}
\end{figure}

@@

The high-level description of Demes models is focused on the efficiency of
human understanding and the avoidance of errors. We have adopted the widely
used YAML format~\citep{ben2009yaml} as the recommended means of interchanging
Demes models (see Figure~\ref{fig-example} for examples). YAML is a data
serialisation language with an emphasis on simplicity, and interoperates well
with JSON (indeed, YAML 1.2 is a superset of JSON). We chose YAML over JSON
because although JSON is an excellent format for data interchange, it is
ill-suited for human understanding or manipulation. We also considered other
declarative data exchange formats such as TOML,
but chose YAML because of equivalance with JSON,
its popularity and good software support.
Since the Demes data model is defined in JSON Schema,
however, there is no formal dependency on YAML and implementations may choose
to use JSON directly if they wish (e.g., for greater efficiency).

@@

The Demes specification is by design simple and static---we wish to
unambiguously describe a demographic model with fully specified
and fixed parameters. This simplicity means that we cannot directly
specify distributions or estimated confidence intervals
for those parameters. While it is not difficult to imagine extending
the specification in ways that would allows us to do this, it would
greatly increase the complexity of parsers and would likely be of
little practical use in any case. This is because the parameters of
demographic models are typically tightly coupled, and cases in which
distributions for different parameters can be simply described are rare.
In this situation, the simplest way to describe estimated
distribution is simply to list a large number of samples from
the posterior. While writing out a large number of Demes models in
YAML format may seem inefficient, it can in fact be a compact
way to describe these distributions.
\jkcomment{TODO: add in a quick example here where we take an IM model
and simulate a posterior distribution under some reasonable model. Report
how much space it takes to represent a large number of samples from this
distribution}

Similarly, one might be interested in running simulations in which
the demographic model parameters are drawn from a distribution, e.g.,
in ABC inference~\citep{beaumont2002approximate}. Allowing for
the many distributions of interest would greatly increase the complexity of
parsers, with a relatively limited benefit. It would be far simpler
and more powerful to write some code to sample directly from the
distributions of interest, using, for example the Demes Python
API.
\jkcomment{This is garbled, but we get the point? You're probably never
going to be doing ABC from a completely general Demes model as input.
There will be some basic model that you have in mind, that you're generating
distributions of parameters around. So, what's the point in writing
general machinery to do this sampling, when you could write it yourself in
10 lines of Python?}
\jkcomment{Key point: if there was a concise way of writing down the joint
distribution of all the parameters, we wouldn't be doing simulations in the
first place.}
